%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                          %
% Copyright (c) 2018 eBay Inc.                                             %
% Modifications copyright (c) 2018-2019 Anders Berkeman                    %
%                                                                          %
% Licensed under the Apache License, Version 2.0 (the "License");          %
% you may not use this file except in compliance with the License.         %
% You may obtain a copy of the License at                                  %
%                                                                          %
%  http://www.apache.org/licenses/LICENSE-2.0                              %
%                                                                          %
% Unless required by applicable law or agreed to in writing, software      %
% distributed under the License is distributed on an "AS IS" BASIS,        %
% WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied. %
% See the License for the specific language governing permissions and      %
% limitations under the License.                                           %
%                                                                          %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\label{chap:urd}

This chapter is the continuation of chapter~\ref{chap:urd_basic},
``Basic Build Scripting''.  Please read about builds script and
joblists before proceeding.



\section{Introduction to Urd}

Urd is a transaction log-based database and server used to group, tag,
and keep track of jobs.  Build scripts are capable of a large variety
of complicated tasks, and on top of that Urd adds the capabilities of
job organisation, storage, and retrieval.  This makes it possible to
handle much larger and more advanced projects while the process is
fully transparent and reproducible.

Using Urd, a project can be separated into functionally independent
parts, or \textsl{sessions}, each containing at least one job, and all
dependencies between jobs inside as well as between these sessions are
tracked by a \textsl{transaction log}.  Sufficient information is
stored so that any part or session of a project could theoretically be
re-constructed the way it looked at any instance in time.  More
formally, Urd provides two things:
\begin{itemize}
\item[1.] Separation between build scripts,
          and a way to share information about existing sessions
          between different scripts (or the same script at different
          points in time).
\item[2.] A searchable transaction log database of all sessions,
          together with their dependencies on other session.  A
          timestamp, date, integer, or combination thereof can be used
          as key.
\end{itemize}
Due to its transaction log, there is no way using the provided API to
accidentally modify or destroy data stored in the Urd database.  The
interesting transaction log database and many other aspects of Urd
will be explained in this chapter.





\section{A Simple Use Case}

A simple use case is presented here to illustrate the basic Urd
features.  Assume a project where movie recommendation data is to be
analysed.  Every hour, new data corresponding to the last hour's
activity is added to the project in the shape of a log file.  The
project is based on two build scripts, one \textsl{import} script, and
one \textsl{analysis} script.

\subsection{The Import Build Script}
The import scrip is used to look for new data files, and import and
chain them as they become available.  For each new file that is
imported, the build script will tell the Urd server
the \textsl{timestamp} of the file as well as a \textsl{list of all
jobs}, i.e.\ the Urd session, that have been built based on that file.
Typically, this includes \texttt{csvimport}, \texttt{dataset\_type},
perhaps a \texttt{dataset\_hash\_partition}, and similar data
pre-processing jobs.



\subsection{The Analysis Build Script}
The analysis build scrip is used for the data analysis work, and is
perhaps run less regularly.  This script needs to know which jobs that
correspond for the most recently imported file, and this is a
straightforward thing to ask Urd.  When all processing is done, all
the analysis jobs are also stored in Urd together with a corresponding
timestamp so that they can be retrieved later.



\subsection{Urd Provides Separation}
In the example, Urd is used to forward information about executed jobs
from the first build script (the import script) to the second (the
analysis script).  In this sense, Urd provides \textsl{isolation} by
``message passing'' between build scripts.  The import scripts can be
modified or even removed, but as long as no workdirs are cleared and
the Urd database is intact, it is still possible to fetch references
to the existing import jobs from Urd.  Thus, it is always possible to
go back and see what things looked like at any previous point in time.



\subsection{Transparency}
Urd also provides \textsl{transparency}, since it can tell which input
files that were used by any analysis job.  In this way, it will never
happen that there is uncertainty about which data that was used for a
particular analysis.

To investigate the result of an analysis, Urd can be asked to return
the jobs corresponding to, say, the most recent analysis jobs.  The
query response from Urd will contain these jobs, \textsl{as well as}
the results from any Urd queries that were carried out for the jobs to
run.  This information will contain the import jobs, since these were
queried in order to perform the analysis.





\section{Local or Shared Urd Server}
By default, Urd is run as a \textsl{local} server, which means that it
is not externally accessible.  A non-local Urd server that can share
information between several users is straightforward to set up, see
section~\ref{sec:urd_setup}.





\section{Urd Sessions and Lists}
A simple file import script will be used as example in this section:
\begin{python}
def main(urd):
    urd.build('csvimport', filename='txn1.txt')
\end{python}
In order to use this import job in a future context,
a \textsl{session} is created by wrapping the code by
the \texttt{urd.begin()} and \texttt{urd.finish()} functions, like this
\begin{python}
def main(urd):
    urd.begin('import/txn', '2018-05-03')
    urd.build('csvimport', filename='txn1.txt')
    urd.finish('import/txn')
\end{python}
Everything that happens between \texttt{.begin()} and \texttt{.finish()}
makes up the session.  The \texttt{.finish()} function makes sure that
the session is stored permanently to disk for future reference.  In
this case, the session can be retrieved knowing its \textsl{list} name
\begin{verbatim}
    import/txn
\end{verbatim}
and its \textsl{timestamp}
\begin{verbatim}
    2018-05-03
\end{verbatim}
The list identifier is composed of two parts, \texttt{<user>/<list>}, where
\texttt{<user>} is for authorisation purposes.  Each user can
have any number of lists that are globally readable, but only the
authorised user can write to them, as will be explained later.



\section{A First Urd Query}
The list created in the previous section is stored in the Urd database
and therefore ready to be used by other build scripts.  For example,
here is a build script that does some processing on the previously
imported file
\begin{python}
def main(urd):
    urd.begin('process/test')

    import_session = urd.latest('import/txn')

    import_timestamp = import_session.timestamp
    import_job       = import_session.joblist['csvimport']

    urd.build('process', source=import_job)

    urd.finish('process/test', import_timestamp)
\end{python}
The first thing that happens is that all processing is covered in a
session named \texttt{process/text}.  The timestamp can be set
in \textsl{either} \texttt{.begin()} or \texttt{.finish()}.

The script is then retrieving the most recently created urd session
stored in list \texttt{import/txn}.  Two things are extracted from
this data, the \textsl{timestamp} and the \textsl{joblist}.  The
timestamp will be used for this session as well, to indicate that
processing is based on data with that particular timestamp.  A
reference to the \texttt{csvimport} job is then extracted from the
joblist and fed to the \texttt{process} job as an input dataset
parameter. (For information about the \texttt{joblist} class, see
section~\ref{sec:joblist}.)



\section{The Contents of a Stored Session}
\label{sec:urd_item}
Calling \texttt{urd.finish()} will update the Urd database with the
contents of the current \textsl{session}.  Each session is addressable
using a \textsl{list} name (in the format \texttt{<user>/<list>}) and
a \textsl{timestamp}.  Session data is stored internally in the JSON
format, and in build scripts it will appear as a Python \texttt{dict}.
The example presented earlier in this chapter may have been recorded
similarly to this
\begin{json}
{
    "user": "processing",
    "automata": "test",
    "timestamp": "2018-05-03",
    "caption": "",
    "joblist": [
        [
            "process",
            "TEST-37"
        ]
    ],
    "deps": {
        "import/txn": {
            "timestamp": "2018-05-03",
            "caption": "",
            "joblist": [
                [
                    "csvimport",
                    "TEST-34"
                ]
            ],
        }
    },
}
\end{json}

\noindent (This example states that at timestamp \texttt{2018-05-03} in list \texttt{processing/test},
there exists a \texttt{process} job \texttt{TEST-34} that
used a \texttt{csvimport} job \texttt{TEST-34}.  This job
also exists in the urd list \texttt{import/txn} at
timestamp \texttt{2018-05-03}.)

The most important keys are
\starttabletwo
\RPtwo \texttt{timestamp} & Timestamp of session\\[1ex]
\RPtwo \texttt{caption} & A caption\\[1ex]
\RPtwo \texttt{user/automata} & Name of Urd list\\[1ex]
\RPtwo \texttt{joblist} & An object of type \texttt{joblist}, containing all jobs built in the session.  For more information, see~\ref{sec:joblist}.\\[1ex]
\RPtwo \texttt{deps} & A dictionary of dependencies from \texttt{user/automata} to urd sessions: \mintinline{python}@{'user/automata': session}@.\\[1ex]
\stoptabletwo





\section{Urd Sessions:  \texttt{begin()} and \texttt{finish()}}

There are a number of options associated with a session, as shown
here,
\begin{python}
urd.begin(urdlist, timestamp, caption=None, update=False)
urd.finish(urdlist, timestamp, caption=None)
\end{python}
and the following applies
\starttabletwo
\RPtwo \texttt{urdlist} & is the name of the Urd list, and the same
  \texttt{urdlist} must be specified in both \texttt{begin()}
  and \texttt{finish()}.

  The \texttt{urdlist} is specified as \texttt{<user>/<list>}, where
  the \texttt{<user/>} part is optional.  The \texttt{user} string
  is also for authentication, and must correspond to the
  current \texttt{URD\_AUTH} settings, see section~\ref{sec:urd_setup}.\\[2ex]

\RPtwo \texttt{timestamp} & is \textsl{mandatory}, but could be set in either
  \texttt{begin()}, \texttt{finish()}, or both.  \texttt{finish()}
  will override \texttt{begin()}.\\[2ex]

\RPtwo \texttt{caption} & is \textsl{optional}, and can be set in either
  \texttt{begin()} or \texttt{finish()}.  \texttt{finish()}
  will override \texttt{begin()}.\\[2ex]

\RPtwo \texttt{update} &  If set to \pyTrue, the last item in the list may be updated.
  This option will be discussed in section~\ref{sec:trunc-update}.\\[2ex]
\stoptabletwo

The Urd transaction database will be written to only when
the \texttt{.finish()} function is called.  Nothing is stored before
calling \texttt{.finish()}, and it is perfectly okay to
omit \texttt{finish()} to avoid storage or during development work.



\subsection{What if a Build Script is Run Again?}
The Urd database is based on transaction log files.  There is no way
to modify or remove information in the database using the Urd API.
The first time a build script is run, it will cause jobs to be built,
and if a session is set up using \texttt{begin()} and \texttt{finish},
this session will be stored in the Urd database.  The second time the
same script is run, the Accelerator will look up already built jobs
and immediately return job references instead of building anything.
As long as there are no changes, the session will look identical to
the one stored in the Urd database, and it is possible to
execute \texttt{finish()} without errors.  But if there are any
discrepancies, such as a job being rebuilt, Urd will refuse to store
the differing session and instead cause an exception to notify the
user.

Normally, a build script can be written in such a way that re-running
it will be consistent will the Urd database and everything is fine.  A
mismatch with Urd is then an indication of some error.  But there are
cases when re-writing Urd history is the desired option, and this will
be discussed in section~\ref{sec:trunc-update}.





\subsection{Timestamp Definition and Resolution}
The ``timestamp'' used to access items may be stated as either
a \texttt{date}, \texttt{datetime}, \texttt{"datetime"}, \texttt{int}, \texttt{(datetime,
int)}, or \texttt{"datetime+int"}.  Here, \texttt{"datetime"} is a
string of format
\begin{shell}
'%Y-%m-%d %H:%M:%S.%f'
\end{shell}
(See Python's \texttt{datetime} module for explanation.)  A specific
timestamp could be shorter than the above specification in order to
cover wider time ranges.  The following examples cover all possible
cases.
\begin{python}
'2016-10-25'                 # day resolution
'2016-10-25 15'              # hour resolution
'2016-10-25 15:25'           # minute resolution
'2016-10-25 15:25:00'        # second resolution
'2016-10-25 15:25:00.123456' # microsecond resolution
\end{python}
Example of a timestamp with an \texttt{int}
\begin{python}
'2016-10-25+3'
\end{python}
Note that
\begin{itemize}
  \item[--] \texttt{int}s sorts \textsl{first}, 
  \item[--] \texttt{datetime}s without \texttt{int} sorts \textsl{before} \texttt{datetime}s with \texttt{int}s, 
  \item[--] shorter \texttt{datetime} strings sorts \textsl{before} longer \texttt{datetime} strings, and
  \item[--] a timestamp must be > 0.
\end{itemize}





\section{Retrieving an Urd Session}
A specific session can be retrieved from the Urd database using its
\textsl{key} and \textsl{timestamp}.  There are two sets of functions
assigned for this, one that will \textsl{record} the lookup to
the \textsl{ongoing} session, and one that will not.  The reason for
recording lookups is to provide the possibility for a session to
contain references to all other sessions used to create it, and
thereby ensuring transparency.

This section is devoted to the function calls that are recorded, and
these are \texttt{get()}, \texttt{first()}, and \texttt{latest()}.
For any of these calls to work, they have to be issued
from \textsl{within} a session, i.e.\ after a \texttt{begin()} call.
Otherwise Urd would not be able to record session dependencies.
Section~\ref{sec:avoid_recording_urd_deps} describes the non-recording
versions, and more ways to find sessions and inspect Urd database
contents is described in section~\ref{sec:more_urd_search}.



\subsection{Finding an Exact or Closest Match:  \texttt{get()}}
\label{sec:urd_get}

The \texttt{get()} function will return the single session, if
available, corresponding to a specified list and timestamp, see the
following example
\begin{python}
urd.begin('ab/anotherlist')
urd.get("ab/test", "2018-01-01T23")
\end{python}
The timestamp must match exactly for an item to be returned.  If there
is no matching item, the \texttt{get()}-call will return
an \textsl{empty session}, i.e.\ something like this
\begin{python}
{'deps': {}, 'joblist': JobList([]), 'caption': '', 'timestamp': '0'}
\end{python}
The strict matching behaviour can be relaxed by prefixing the
timestamp with one of
\begin{itemize}
\item[] ``\texttt{<}'', ``\texttt{<=}'', ``\texttt{>}'', or ``\texttt{>=}''.
\end{itemize}
For example
\begin{python}
urd.get("ab/test", ">2018-01-01T01")
\end{python}
may return an item recorded as \texttt{2018-01-01T02}.  Relaxed
comparison is performed ``from left to right'', meaning that
\begin{python}
urd.get("ab/test", ">20")
\end{python}
will match the first recorded session in a year starting with
``\texttt{20}'', while
\begin{python}
urd.get("ab/test", "<=2018-05")
\end{python}
will match the latest timestamp starting with ``\texttt{2018-05}'' or less,
such as ``\texttt{2018-04-01}'' or ``\texttt{2018-05-31T23:59:59.999999}''.



\subsection{Finding the Latest Session:  \texttt{latest()}}
The \texttt{latest()} call will, for a given \textsl{key}, return the
session with most recent timestamp.  For example
\begin{python}
urd.begin('ab/anothertestlist')
urd.latest('ab/test')
\end{python}
will return a complete session, assuming that the list is exists and
is populated.  Otherwise, an \textsl{empty session} will be returned.



\subsection{Finding the first item:  \texttt{first()}}
The \texttt{first()} function works similarly to \texttt{latest}, but
will instead return the session with the \textsl{oldest} timestamp,
i.e.\ the first session stored using the \textsl{key}.





\section{Retrieving an Urd Session without Recording}
\label{sec:avoid_recording_urd_deps}
As pointed out in the previous section, dependencies will be recorded
when the \texttt{get()}, \texttt{latest()}, and \texttt{first}
functions are used.  If, for some reason, the point is to just have a
look at the database to see for example if something is already there
and in that case what that is, it can be done using the
``\textsl{peek}'' functions: \texttt{peek()}, \texttt{peek\_first()},
and
\texttt{peek\_latest()}, like this:
\begin{python}
urd.peek('test', '2016-10-25')
urd.peek_latest('test')
urd.peek_first('test')
\end{python}
There are some good use cases for this, but it is not recommended for
general use, they may cause a loss of continuity and visibility.





\section{Aborting an Urd Session:  \texttt{abort()}}

There can only be one ongoing Urd session.  A new session cannot be
started until the current one has finished.  There are three ways to
end an urd session
\begin{itemize}
\item[-] execute the \texttt{finish()} call and have the session recorded.
(Or rejected, if a session with the same key and timestamp already
exists but with different contents.)
\item[-] end the build script ``prematurely'' before the
\texttt{finish()}-call.  No data will be stored in Urd.
\item[-] issue an \texttt{abort()} call.
\end{itemize}
The \texttt{abort()} function is used like this
\begin{python}
urd.begin('test')
urd.abort()
# execution continues here, a new session can be initiated
urd.begin('newtest')
\end{python}
Similar to unfinished sessions, aborted sessions will not be stored in
the Urd transaction log.





\section{Truncating and Updating}
\label{sec:trunc-update}
Since the Urd database is based on transaction log files, it will
always keep a consistent history of all events taken place.  It is not
possible to erase or modify old entries, but it is okay to update the
latest item, or set a marker in the log indicating that the list is
starting over from a certain date and everything before this marker
should not be considered anymore.  This makes it possible to both
keeping the full history \textsl{and} being able to rewrite it.  There
is full transparency and reproducibility -- all sessions before an
update or restart marker are always kept in the Urd log file,
transaction data is always \textsl{appended} to the log files.



\subsection{Updating the Last Item}
To update the last item in a list, set the \texttt{update} argument
to \pyTrue
\begin{python}
urd.begin('test', '2016-10-25', update=True)
\end{python}
If update is \pyTrue, the entry in the test list at '2016-10-25' will
be updated, unless the new information is equivalent.
The \texttt{update()} call will simply add a new line to the Urd log
database, and if the timstamp is the same as the previous entry, the
new entry will be selected.  Note that this only works for
the \textsl{last} entry in the database, and that the previous version
is not erased, it still exists in the log file.



\subsection{Truncating a List}
Sometimes there is a need to wipe all history or at least start over
from an earlier time instance.  Since data can not be erased from a
transaction log, re-winding to an earlier timestamp is implemented
using inserting a ``truncation'' marker in the log.  Any data more
recent than what is indicated by the marker is wiped from memory, and
sessions with timestamps more recent than the marker can be stored
again.  The \texttt{truncate()} function is used to insert these
markers, and it is used like this
\begin{python}
urd.truncate(ab/'test', '2016-09-30')
\end{python}
This will rollback everything that has happened in
the \texttt{ab/test} list back to '2016-09-30'.  There is also a
special case,
\begin{python}
urd.truncate('ab/test', 0)
\end{python}
that will erase all items from memory and cause the list to start over
again.  Again, remember that Urd is using plain text transaction log
that can only be appended to.  It is always possible to recover any
old result or processing state from these logs.



\subsection{Truncation Consequences:  Ghosts}
When a list is truncated, all items after a specified timestamp are
made invisible.  Assuming that another list has stored a dependency of
an item that is truncated, the jobs in this list are now without
dependencies that can be looked up.  We call them ``ghosts''.  Ghosts
cannot be looked up in Urd, but they are still in the database, where
they are marked as ghosts.





\section{More Search Functions}
\label{sec:more_urd_search}

There are two more functions for finding information in the Urd
database: \texttt{list} and \texttt{since}.



\subsection{Listing all Urd Lists:  \texttt{list()}}
The \texttt{list()} function will return a list of all lists recorded in
the database:
\begin{python}
print(urd.list())
\end{python}
may show something like
\begin{shell}
['ab/test', 'ab/live', 'production/import', 'production/live']
\end{shell}



\subsection{Listing all Items After a Specific Timestamp:  \texttt{since()}}
The \texttt{since()} function is used to extract lists
of \textsl{timestamps} corresponding to recorded sessions.  In its
most basic form, it is called with a timestamp like this
\begin{python}
urd.since('ab/test', '2016-10-05')
\end{python}
which returns a list with all existing timestamps more recent than the
one provided
\begin{shell}
['2016-10-06', '2016-10-07', '2016-10-08', '2016-10-09', '2016-10-09T20']
\end{shell}
The \texttt{since()} function is rather relaxed with respect to the
resolution of the input.  The input timestamp may be truncated from
the right down to only one digits.  An input of zero is also valid.
For example, these are all valid
\begin{python}
urd.since('ab/test', '0')
urd.since('ab/test', '2016')
urd.since('ab/test', '2016-1')
urd.since('ab/test', '2016-10-05')
urd.since('ab/test', '2016-10-05T20')
urd.since('ab/test', '2016-10-05T20:00:00')
\end{python}





\section{Building Jobs: \texttt{build()}}
\label{sec:urd_build}

Jobs are dispatched in Urd sessions using the \texttt{build} function.
Here is the complete call with all possible parameters.
\begin{python}
job = urd.build(
    method,
    options={},    datasets={},    jobs={},
    name='',       caption='',
    workdir=None,
    **kw
)
\end{python}
If \options, \datasets, and \jobs are uniquely defined in the method,
they could be entered just as plain keyword arguments.  If there are
ambiguities, the full \texttt{options=} etc.\ must be used.  Here is
an explanation of \texttt{build} parameters:
\starttabletwo
\RPtwo \texttt{method} & Name of method to build.  Enter \texttt{test}
    here if the method filename is \texttt{a\_test.py}.  The
    Accelerator will look for this file in all method directories
    specified in the Accelerator's configuration file.\\

\RPtwo \texttt{kw} & Use to specify any unique options, jobs, and datasets
    without the more elaborate dictionaries described next.\\
    
\RPtwo \texttt{options=\{\}} & A \texttt{dict} of options to the method.
    This overrides options defined in the method itself, but adding
    options not prototyped in the method is \textsl{not} allowed.\\

\RPtwo \texttt{jobs=\{\}} & A \texttt{dict} of jobs to the method.
    It is possible to specify a list of jobs like this
\begin{python}
jobs=dict{alljobs=[job1, job2,...]}
\end{python}
\\

\RPtwo \texttt{datasets=\{\}} & A \texttt{dict} of datasets to the method.
    Datasets may be lists too, just like \texttt{jobs} above.\\

\RPtwo \texttt{workdir=None} & If specified, the job will be built in
    this workdir, assuming the workdir is specified in the
    configuration file as \textsl{either} source or target.\\

\RPtwo \texttt{name} & A string associated with the job.  Use it
    to distinguish several jobs created from the same method in
    joblists.\\
    
    \texttt{caption} & A caption string.  For decorative purposes
    only, this has no practical use.\\
\stoptabletwo

The \texttt{build()} function will only build a job when it has to,
otherwise it will just return a job reference to an existing matching
job.  In order to match, an existing job must have
\begin{itemize}
\item[-] exactly the same source code, i.e.\ the \textsl{hash} of the source code must match,
\item[-] exactly the same options, datasets, and jobs.
\end{itemize}
If the source code is changed, a job rebuild can be prevented using
the \texttt{equivalent\_hashes} variable as explained in
section~\ref{sec:equivalent_hashes}.



\subsection{Building Chained Jobs: \texttt{urd.build\_chained()}}
This is a special version of \texttt{build()} that can be used for
linking a set of dataset-creating jobs.  This function was created for
the purpose of having build scripts that imported a large set of files
in a \texttt{for}-loop.  It is used like this
\begin{python}
def main(urd):
    # Import a list of files
    for filename, timestamp in listoffiles:
        urd.begin('import', timestamp)
        urd.latest('import')
        job = urd.build_chained('csvimport',
                                filename=filename,
                                ...
                                name='importing')
        job = urd.build_chained('dataset_type',
                                source=job,
                                ...
                                name='typing')
        urd.finish('import')
\end{python}
This example will build a chain of \texttt{csvimport} jobs, and one
chain of \texttt{dataset\_type} jobs.  Each \texttt{dataset\_type} job
will have a corresponding \texttt{csvimport} dataset as source.
The \texttt{build\_chained()} function works, provided that
\begin{itemize}
\item[--] the method to build has a dataset named \texttt{previous},
\item[--] a unique \texttt{name=} is set in \texttt{.build\_chained()}, and
\item[--] \texttt{urd.latest()} is called inside the Urd session.
\end{itemize}
The call to \texttt{urd.latest()} is necessary for the
dependency-logic to work, but the output from the call can be
discarded.





\section{Changing workdir:  \texttt{set\_workdir()}}
\label{sec:set_workdir}
The target workdir specified in the configuration file is the only
workdir that is written to by default.  Any other workdir is read
only.  This behaviour can be overridden, either
\begin{itemize}
\item[] per job, using the \texttt{workdir=...} option to \texttt{urd.build} as shown in section~\ref{sec:urd_build}, or
\item[] using \texttt{urd.set\_workdir()}.
\end{itemize}
The latter,
\begin{python}
def main(urd):
    urd.set_workdir(<workdir>)}
\end{python}
will set the workdir for all coming \texttt{build} calls in the
current build script.  It can still be overridden using
the \texttt{workdir=} option to \texttt{urd.build}.





\section{Profiling a Build Script:  \texttt{print\_exectimes()}}
The \texttt{JobList} object has a helper function that can be used to
print profiling information for the joblist.  The following example is
self-explanatory
\begin{python}
def main(urd):
    ...
    urd.joblist.print_exectimes()
\end{python}
This will print execution times for all jobs in the session
to \texttt{stdout}.  It may for example look like this
\begin{verbatim}
Time per method:
   color2        23.7 seconds  (25%)
   csvexport     17.5 seconds  (18%)
   lowpass2      15.6 seconds  (17%)
   newcol        14.4 seconds  (15%)
   black          5.7 seconds  (6%)
   colimage       5.4 seconds  (6%)
   sync           4.7 seconds  (5%)
   clamp          3.8 seconds  (4%)
   dataset_type   2.7 seconds  (3%)
   csvimport      1.4 seconds  (1%)
Total time 94.8 seconds
\end{verbatim}
The methods are sorted by execution time, top to bottom.





\section{Passing Flags from the Command Line}
Flags can be added to a build script at run time using a comma
separated list like this
\begin{python}
ax run [script] --flags=verbose,skiptest
\end{python}
The flags will appear in the \texttt{urd}-object like this
\begin{python}
def main(urd):
    if 'verbose' in urd.flags:
       print('verbosity')
\end{python}





\section{The Urd HTTP-API}
Urd can be accessed directly without using the Accelerator by calling
its HTTP API.  These calls are easy to use and adds transparency since
the database contents can be peeked from the command line without
writing any programs.  Here is a list of all API endpoints
\begin{verbatim}
  list
  <user>/<list>/first
  <user>/<list>/latest
  <user>/<list><timestamp>
  <user>/<list>/since/<timestamp>
\end{verbatim}
All calls return data in the JSON format.  The Accelerator comes with
a built in ``\texttt{curl}'' command that is designed for these calls,
and it is used like this
\begin{shell}
% ax curl <endpoint>
\end{shell}
But any standard tool such as the famous \texttt{curl} program works
too, for example like this
\begin{shell}
% curl http://localhost:8123/list
\end{shell}



\subsection{The \texttt{list} endpoint}
To show all stored lists issue
\begin{shell}
% ax curl list
["ab/test"]
\end{shell}
All available lists are returned in a typed JSON list.



\subsection{The \texttt{since} endpoint}
The \texttt{since} endpoint is used to get a list of all entries more
recent than a timestamp.  For example, to see information added
after \texttt{2016-10-24}, do
\begin{shell}
% ax curl ab/test/since/2016-10-24
["2016-10-25"]
\end{shell}
\begin{shell}
% ax curl ab/test/since/2016-10-26
[]
\end{shell}
Results are in JSON list format.



\subsection{The \texttt{first} and \texttt{latest} endpoints}
Looking up the latest stored job in the \texttt{ab/test} list
\begin{shell}
% ax curl ab/test/latest
{"caption": "", "automata": "test", "user": "ab", "deps": {},
  "timestamp": "2016-10-25", "joblist": [["method1", "test-56"],
  ["method2", "test-59"], ["method3", "test-60"]]}
\end{shell}
And see the first stored job in the test list
\begin{shell}
% ax curl ab/test/first
\end{shell}
works similarly.  The returned data is an Urd item, described in
section~\ref{sec:urd_item}, in JSON format.



\subsection{The \texttt{get} endpoint}
Actually, there is no explicit \texttt{get} endpoint.  Instead, the
API is just called by the name of the list and a timestamp.  For
example, to see what is inside the test list stored
at \texttt{2016-10-25}
\begin{shell}
% ax curl ab/test/2016-10-25
{"caption": "", "automata": "test", "user": "ab", "deps": {},
  "timestamp": "2016-10-25", "joblist": [["method1", "test-56"],
  ["method2", "test-59"], ["method3", "test-60"]]}
\end{shell}
The timestamp may be truncated to the right, and prefixed
by \texttt{>}, \texttt{>=}, \texttt{<}, and \texttt{<=}, just as
described in section~\ref{sec:urd_get}.  \textsl{Make sure to quote
the request if these characters are used in a call from the shell.}





\section{Multi-Used Urd Consistency}
Urd can be accessed by a large number of clients.  Each client may add
to or truncate any list at any time.  In order to avoid race
conditions and make the database deterministic, all \texttt{add}-
and \texttt{truncate}-requests appears in a sequential manner to the
Urd server.  Each request is assigned with an unique timestamp, and
stored in the requested list.

When Urd is restarted, it reads all the database files, and sorts all
rows in order of the receive timestamp.  Thereafter, each row is
applied in increasing time order to the internal, RAM-based database.
Due to the unique timestamping, the result is a deterministic replica
of the previous run.
