%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                          %
% Copyright (c) 2018 eBay Inc.                                             %
% Modifications copyright (c) 2019-2020 Anders Berkeman                    %
%                                                                          %
% Licensed under the Apache License, Version 2.0 (the "License");          %
% you may not use this file except in compliance with the License.         %
% You may obtain a copy of the License at                                  %
%                                                                          %
%  http://www.apache.org/licenses/LICENSE-2.0                              %
%                                                                          %
% Unless required by applicable law or agreed to in writing, software      %
% distributed under the License is distributed on an "AS IS" BASIS,        %
% WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied. %
% See the License for the specific language governing permissions and      %
% limitations under the License.                                           %
%                                                                          %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\label{chap:urd_basic}

Build scripts are used to execute jobs and control the job flow on the
Accelerator.  This chapter describes the basics of job building.  More
advanced features, using the Urd server, are presented in
chapter~\ref{chap:urd}.



\section{Build Scripts}
\index{build scripts}
Build scripts are stored in \textsl{method directories}.  The default
build script is named ``\texttt{build.py}'', all other build scripts
have to start with the string ``\texttt{build\_}''.  Build scripts are
executed using the \texttt{run} command.  For example,
\begin{shell}
ax run example
\end{shell}
will look for a file named \texttt{build\_example.py} and execute it,
while the command
\begin{shell}
ax run
\end{shell}
by itself will execute the default build script.  (The \texttt{run}
command is described in section~\ref{sec:exec_runner}.)

The \texttt{run} command will load the build script and execute
its \texttt{main()} function.  This function takes
a \textsl{mandatory} argument named \texttt{urd}, so a basic build
script looks like this
\index{build script!\texttt{main} function}
\begin{python}
def main(urd):
    # do something here...
\end{python}
The \texttt{run} command inserts an object of the \texttt{Urd} class
as the argument to the \texttt{main()} function.  This \texttt{urd}
object has a number of member functions and attributes useful for job
building and tracking of old jobs.  The \texttt{Urd} class is
described in chapter~\ref{chap:urd} and in
section~\ref{sec:classes:urd}.



\subsection{Building a Job: \texttt{urd.build()}}
The \texttt{.build()} function is used to build a job from a method
(i.e.\ source file).  For example, the most simple build script that
executes a method, here called \texttt{method1}, is
\begin{python}
def main(urd):
    urd.build('method1')
\end{python}
Common options for the \texttt{build} function are as follows
\index{\texttt{urd.build}}
\begin{python}
job = urd.build(method,
                options={}, datasets={}, jobs={},
                name='', caption='', workdir=None, **kw)
\end{python}
In practive, only a few parameters are given at a time, and the only
mandatory argument is the \texttt{method}.  Any arguments to a job
build can be added as simple \textsl{keyword arguments}
(\mintinline{python}{key=value}), and the build process will
automatically assign them by name to the corresponding
\texttt{options}, \texttt{datasets}, or \texttt{jobs} parameter
sets, assuming the naming is unique.  If the key name is not unique,
the argument has to be specified explicitly.  Examples will be present
in the following sections.

When the job is completed, Urd will record it using the name of the
method as key, unless the \texttt{name=} is specified.
The \texttt{name} parameter is particularly useful to tell jobs apart
that are based on the same method.  A common case would be
the \texttt{csvimport} method, for example.  It is also possible to
assign a caption to a job, but this has no functional benefits.

When the job has been successfully built, the \texttt{build} function
will return a reference of type \texttt{Job}.  The \texttt{Job} class
contains member functions and attributes that can be used to extract
information, such as generated files or text written
to \texttt{stdout}, from the job.  The \texttt{Job} class is described
in detail in chapter~\ref{chap:jobs} and in section~\ref{sec:classes:job}.

Similarly, if the job to be built already exists in a
configured \textsl{workdir}, the \texttt{build} function will
immediately return a \texttt{Job} object corresponding to the existing
job without executing anything.



\subsection{Connecting Jobs}
Jobs may be connected, or linked, together so that a new job may use
the output generated by previous jobs.  In this way, jobs could share
and re-use previous results, and complicated tasks can be broken down
into several smaller jobs.  A link is created by feeding the job
object from an existing job into the \texttt{.build()}-call of a new
job, like in this example
\begin{python}
def main(urd):
    job_import  = urd.build('csvimport', filename='inputfile.txt')
    job_process = urd.build('process',   source=job_import)
\end{python}
In the example above, the first job, \texttt{csvimport}, imports the
file ``\texttt{inputfile.txt}''.  The second job, \texttt{process},
takes the imported dataset as input for further processing.



\subsection{Replaying Build Scripts}

When the example build script from the previous section is run, both
the \texttt{csvimport} and the \texttt{process} jobs will be built.
But what happens if the same build script is run a second time?
Remember now that the Accelerator stores all jobs in associated
workdir(s).  If there has been no changes to the code since the last
run, the Accelerator will immediately find the job reference to
the \texttt{csvimport} without needing to execute it again.  The
reference will be input to the \texttt{.build()} call of the second
method, and since the Accelerator has seen this call before too, it
will immediately look up the reference to this job as well, instead of
executing it.  A second run of the build script will only take a
fraction of a second to execute, but it will still return all job
references.

On the other hand, if something has been modified, such as a method's
source code or any of the input parameters, the affected job(s) will
be re-executed.  For example, assume that the \texttt{csvimport} build
call is modified to import ``\texttt{anotherfile.txt}'' instead.  This
will cause the \texttt{csvimport} method to be executed again, leading
to a new job with a new jobid.  This job reference is input to
the \texttt{process} job, causing it to be re-executed too.  Here,
both jobs were re-executed, since they both were affected by the
modification, but in general, only those jobs affected by the
modification will be re-executed!

\textsl{A successful ``replay'' of a build script ensures the integrity and
dependencies of all involved calculations.}  If there are no changes,
the same result remains.  If, however, some of the code has been
modified, the Accelerator will compute new jobs to reflect the new
situation.  The result may be different, and the user is notified.


\section{Working with Build History:  \texttt{urd.joblist}}
\label{sec:joblist}

Information about previously executed jobs is stored in
the \texttt{urd.joblist} variable.  This variable is of
type \texttt{JobList}, which is basically a standard ordered
Python \texttt{list} with some additional features for searching,
profiling and pretty-printing.  The \texttt{JobList} class is further
explained in section~\ref{sec:classes:joblist}.



\subsection{Printing a \texttt{JobList}:  \texttt{urd.joblist.pretty}}
Create a \texttt{JobList} and pretty-print it
\begin{python}
def main(urd):
    job1 = urd.build('first')
    job2 = urd.build('second', first=job1)
    print(urd.joblist.pretty)
\end{python}
which results in
\begin{shell}
JobList(
   [  0]  first : TEST-38
   [  1] second : TEST-39
)
\end{shell}
(The actual jobids will most likely be different.)  The name in
the \texttt{joblist} is either the name of the method, or, if present,
the name given explicitly using the \texttt{urd.build(name=)} option.



\subsection{Finding Jobs in a Joblist}
There are several ways to extract jobs or list of jobs from
a \texttt{joblist}.


\subsubsection*{Using \texttt{find} to Extract Jobs to a New Joblist}
The \texttt{find()} function finds matching jobs and returns them in a
new \texttt{JobList}.  For example,
\begin{python}
jl = urd.joblist.find('csvimport')
\end{python}
will create a \texttt{JobList} of all \texttt{csvimport} jobs
in \texttt{urd.joblist}.


\subsubsection*{Using \texttt{get} to Potentially Extract a Single Job}
The \texttt{get} function will return a job reference to the most
recent matching job.  For example,
\begin{python}
job = urd.joblist.get('csvimport')
\end{python}
If no matching job is found, \texttt{get} will return \pyNone.


\subsubsection*{Using Square Brackets to Extract a Singe Job}
Accessing jobs directly with a key like this
\begin{python}
job = urd.joblist['csvimport']
\end{python}
is similar to \texttt{get}, but will return an error if a matching job
is not found.

%% \subsection{Return a \texttt{JobList} as a \texttt{tuple}}
%% The \texttt{as\_tuples} function will return the joblist as a list of
%% tuples,
%% \begin{python}
%% x = urd.joblist.as_tuples
%% \end{python}
%% will return something like
%% \begin{verbatim}
%% [('csvimport', 'test-0'), ...]
%% \end{verbatim}


\subsection{Indexing and Slicing a \texttt{JobList}}
Since the \texttt{JobList} class is an extension of the \texttt(list)
class, it is possible to do some \texttt{list}-like operations on
joblist objects.

A common special case is to provide an integer to the \texttt{.get()}
function.  This will cause a lookup of the item with the corresponding
index in the list.  It is often used like this
\begin{python}
latest_job = urd.joblist.get(-1)
\end{python}
as a way to return the last item in the joblist, should it exist, or \pyNone.

It is also possible to index and slice like in the following examples
\begin{python}
joblst = jl[3]
\end{python}
or
\begin{python}
joblst = jl[-2:]
\end{python}



\section{Configuration Information:  \texttt{urd.info}}
The dictionary \texttt{urd.info} contains configuration information
from the Accelerator server.  In particular, it contains at least
these fields
\starttabletwo
\texttt{slices} & Configured number of slices.\\
\texttt{urd} & An URL to the Urd server.\\
\texttt{result\_directory} & see section~\ref{sec:configfile}.\\
\texttt{input\_directory} & see section~\ref{sec:configfile}.\\
\stoptabletwo



\section{Summary}
In a build script, the \texttt{urd} object has functionality for
building and retrieving jobs.  A job is built
using \texttt{urd.build()}, and references to all built jobs are
stored in \texttt{urd.joblist}.  These references could be fed as
input parameters to new jobs so that the output from one job could be
used as input by another.  The \texttt{urd.joblist} variable is
basically of type \texttt{list}, but with extra functionality to
conveniently find already existing jobs.
